---
title: "The Sparks Foundation GRIPJUNE Task #4"
author: "Roel George"
date: "09/06/2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
# Exploratory Data Analysis on the Global Terrorism Dataset

## Getting Started  
```{r warning=FALSE, message=FALSE}
# Installing the required packages needed for the Exploratory data anlysis 

#install.packages("ggplot2")
#install.packages("dplyr")
#install.packages("ggmap")
#install.packages("plotly")
#install.packages("pacman")
#install.packages("scales")

# Calling the installed packages
library("scales")
library("plotly")
library("ggplot2")
library("dplyr")
library("ggmap")
library("pacman")
```


## Importing and Cleaning Data.  

```{r warning=FALSE, message=FALSE}

# Reading the Global Terrorism data.
global_terrorism<-read.csv("globalterrorismdb_0718dist.csv")
# Looking at the dimensions of the global_terrorism data
dim(global_terrorism)
# We see that it has 181691 rows and 135 columns, but since most of it is not of much use,
# we create a cleaner data frame consisting of the most usable data.
clean<-global_terrorism[c("iyear","country_txt","city","summary","attacktype1_txt","targtype1_txt","success","gname","weaptype1_txt","propextent_txt","longitude","latitude")]
# Removing the NA values form the clean dataset.
clean<-na.omit(clean)
# Summary and head of the clean data.
summary(clean)
head(clean)

```
### Observation:  
The 'iyear' column shows the year when an terrorism incident occurred. The provided data has recorded global terrorism activities starting form the year 1970 till 2017.  <br/>
The 'country_txt' column shows the country where the incident accurred. We have a total 204 countries in the cleaned dataset.  <br/>
The 'city' column shows the city in which the incident accurred.  <br/>
The 'summary' column shows the short summany(headline) on the incident for few of the rows in the dataset.  <br/>
The 'attacktype1_txt' column shows type of attack which took place.The types of attack are categorized into 9 types.   <br/>
The 'targtype1_txt' column shows the group of individuals on whom the attack was targeted. There are 22 targeted groups in the cleaned dataset.  <br/>
The 'success' column indicates weather the attack was successful or not.  <br/>
The 'weaptype1_txt' column shows the kind of weapons used for the attack.It is classified into 12 groups.  <br/>
The 'gname' column shows the group of people who carried out the act of terrorism.  <br/>
The 'propextent_txt' column shows the amount of property damage(capital) occurred. It is broadly classified into 4 groups.  <br/>  

## Observing the year data according to different factors 
```{r echo=FALSE, warning=FALSE, message=FALSE}

# Creating a box-plot of the year data based on the attack type.  
plot1<-plot_ly(y=clean$iyear,color=clean$attacktype1_txt,type="box",title="Boxlplot of the year data as a factor of Attack type.")
plot1

# Creating a box-plot of the year data based on the target type.  
plot2<-plot_ly(y=clean$iyear,color=clean$targtype1_txt,type="box",title="Boxlplot of the year data as a factor of Target type.")
plot2

# Creating a box-plot of the year data based on the weapon type. 
plot9<-plot_ly(y=clean$iyear,color=clean$weaptype1_txt,type="box",title="Boxlplot of the year data as a factor of Weapon type.")
plot9 %>% layout(showlegend = FALSE)
```  

### Observation:  
We observe how the behavior of different types of global terrorism attacks, using diffent weapon types , targeting different groups of individuals changed over the years.

## A Frequency distribution of acts of terrorism across different parts of the globe.
```{r echo=FALSE}
# Creating a ggplot and storing it in a variable called plot3.
plot3<-ggplot()+
  geom_bar(data=clean,
                 aes(x=(country_txt),fill=as.factor(success)),
                 stat = "count"
                 )+
# Changing the element size of the labels 
  theme(axis.text.x = element_text(angle=90,size =4),
        legend.title = element_text(size=1))+
# Adding title and xlab ad ylab
  labs(
    title = "Country wise distribution of terrorism count",
    x="Country",
    y="frequency"
  )
# Calling the plot3 as a plotly graph.
ggplotly(plot3)

```  

<br/>  

### Observations:  

From this plot we observe that (in the given dataset) the Iraq is the county with maximum number of successful terrorist attacks. Iraq is followed by Pakistan, Afghanistan and India. 

## Observing the trend of golbal terorism from 1970-2017

```{r echo=FALSE}
# Creating a plot and storing it in a variable called plot4.
plot4<-ggplot()+
  geom_line(data=clean,col="#5ab4ac",
                 aes(x=(iyear)),
                 stat = "count")+
  geom_point(
    data=clean,col="#5ab4ac",
                 aes(x=(iyear)),
                 stat = "count"
  )+
# Changing the labels of the plot and aadding the title.
  labs(
    x="year",
    y="count",
    title = "Frequency distriution of the total years."
  )+
  theme(legend.title = element_blank())
# Calling plot4 as a plotly graph.
ggplotly(plot4)

# Creating a plot and storing it in a variable called plot5.
plot5<-ggplot()+
  geom_histogram(data=clean,
                 aes(x=(iyear),fill=as.factor(attacktype1_txt)),
                 stat = "count")+
# Adding facets.
  facet_wrap(as.factor(clean$attacktype1_txt))+
# Removing the legends.
  theme(legend.position = "none")+
# Adding x and y labels and title.
  labs(
    x="year",
    y="count",
    title = "Distribution of various types of attacks throught the years."
  )
# Calling plot5 as a plotly graph
ggplotly(plot5)

```  

<br/>  

### Observations:  

Through the 1st line-plot, we observe that there is a steady increase the number of terrorism activities (note: the data got the year 1993 is missing), followed by a relative decrease an then again followed by a sharp increase till the year 2014.  <br/>
Note: the behavior of the plot depends on multiple factors like world population and the accuracy of the recorded data and hence one shouldn't consider this as te only determining factor.  <br/>

We see that almost a similar trend is observed for the attack type in the 2nd plot in the chunk.  <br/>  

## Observing Percent count of various Attack Type.

```{r echo=FALSE}
# Creating a variable h and storing the plot data (create using ggplot_build()) inside it . 
h<-ggplot()+
  geom_bar(data=clean,
           aes(x=(attacktype1_txt),fill=as.factor(attacktype1_txt)),
           stat = "count",
           col="black",width=1)
h<-ggplot_build(h)
h<-h$data[[1]]
# Adding the percent count of each of the row of the data in a new column.
h$perc<-h$count*100/sum(h$count)
# Adding a column containing the label for each of the groups.
h$label<-levels(as.factor(clean$attacktype1_txt))

# Using the a above create dataset to plot a plotly pie chart.
plot6<-plot_ly( data =h , labels =h$label , values =h$perc , type = "pie", textinfo = "label+percent", insidetextorientation = "radial",title="Pie Chart on the distribution of Attack Type")
plot6

```  

<br/>  

### Observations:  

The most common type of terrorism attack is Bombing/Explosion taking up almost half the pie followed by Armed Assault and Assasination.  

## Observing Percent count of various Target Type.  

```{r echo=FALSE}
# Creating a variable f and storing the plot data (create using ggplot_build()) inside it . 
f<-ggplot()+
  geom_bar(data=clean,
           aes(x=(targtype1_txt),fill=as.factor(targtype1_txt)),
           stat = "count",
           col="black",width=1)
f<-ggplot_build(f)
f<-f$data[[1]]
# Adding the percent count of each of the row of the data in a new column.
f$perc<-f$count*100/sum(f$count)
# Adding a column containing the label for each of the groups.
f$label<-levels(as.factor(clean$targtype1_txt))

# Using the a above create dataset to plot a plotly pie chart.
plot7<-plot_ly( data =f , labels =f$label , values =f$perc , type = "pie", textinfo = "label+percent", insidetextorientation = "radial",title="Pie Chart on the distribution of Target Type")
plot7

 
```  

<br/>  

### Observations:  

Private Citizens and Property is the most commonly targeted group forllowed by the Military, Police,the general Government and Business.  


## A better visualization of the terrorism data across the world.   
```{r echo=FALSE}
# Storing the world map data in a variable called world_map.
world_map<-map_data("world")
# Using the world_map data to plot teh world map using ggplot().
plot8<-ggplot()+
  geom_polygon(data=world_map,aes(x=long,y=lat,group=group),col="black",fill="#80c904")+
  theme(
    panel.background = element_rect(fill = '#59bfff')
  )+
# Using the latitudes and longitudes from the cleaned dataset to add points representing terrorism activities.
  geom_point(data=clean,aes(x=longitude,y=latitude),alpha=0.01)+
  scale_x_continuous(labels = function(x) format(x, scientific = FALSE))+
  coord_cartesian(x=c(-180,180),y=c(-90,90))
# Calling the plotted map.
(plot8)

```  

### Observation:
We observe that there's a high density of clustering around the Middle-East, Southern Asia, the Andean Regions of South America, few parts of Central Africa, Southern Europe, and  South-Eastern Asia.

